### 2018-09-13 paper.tex
VGTC template import

\documentclass{vgtc}                          % final (conference style)
%\documentclass[review]{vgtc}                 % review
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint]{vgtc}               % preprint
%\documentclass[electronic]{vgtc}             % electronic version

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Figures should be in CMYK or Grey scale format, otherwise, colour 
%% shifting may occur during the printing process.

%% These few lines make a distinction between latex and pdflatex calls and they
%% bring in essential packages for graphics and font handling.
%% Note that due to the \DeclareGraphicsExtensions{} call it is no longer necessary
%% to provide the the path and extension of a graphics file:
%% \includegraphics{diamondrule} is completely sufficient.
%%
\ifpdf%                                % if we use pdflatex
  \pdfoutput=1\relax                   % create PDFs from pdfLaTeX
  \pdfcompresslevel=9                  % PDF Compression
  \pdfoptionpdfminorversion=7          % create PDF 1.7
  \ExecuteOptions{pdftex}
  \usepackage{graphicx}                % allow us to embed graphics files
  \DeclareGraphicsExtensions{.pdf,.png,.jpg,.jpeg} % for pdflatex we expect .pdf, .png, or .jpg files
\else%                                 % else we use pure latex
  \ExecuteOptions{dvips}
  \usepackage{graphicx}                % allow us to embed graphics files
  \DeclareGraphicsExtensions{.eps}     % for pure latex we expect eps files
\fi%

%% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
\graphicspath{{figures/}{pictures/}{images/}{./}} % where to search for the images

\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
\usepackage{textcomp}                  % use better special symbols
\usepackage{mathptmx}                  % use matching math font
\usepackage{times}                     % we use Times as the main font
\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
\usepackage{cite}                      % needed to automatically sort the references
\usepackage{tabu}                      % only used for the table example
\usepackage{booktabs}                  % only used for the table example
%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{Global Illumination for Fun and Profit}

%% This is how authors are specified in the conference style

%% Author and Affiliation (single author).
%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}}
%%\affiliation{\scriptsize Allied Widgets Research}

%% Author and Affiliation (multiple authors with single affiliations).
%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com} %
%%\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com} %
%%\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}}
%%\affiliation{\scriptsize Martha Stewart Enterprises \\ Microsoft Research}

%% Author and Affiliation (multiple authors with multiple affiliations)
\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ %
        \scriptsize Starbucks Research %
\and Ed Grimley\thanks{e-mail: ed.grimley@aol.com}\\ %
     \scriptsize Grimley Widgets, Inc. %
\and Martha Stewart\thanks{e-mail: martha.stewart@marthastewart.com}\\ %
     \parbox{1.4in}{\scriptsize \centering Martha Stewart Enterprises \\ Microsoft Research}}

%% A teaser figure can be included as follows, but is not recommended since
%% the space is now taken up by a full width abstract.
%\teaser{
%  \includegraphics[width=1.5in]{sample.eps}
%  \caption{Lookit! Lookit!}
%}

%% Abstract section.
\abstract{Duis autem vel eum iriure dolor in hendrerit in vulputate
velit esse molestie consequat, vel illum dolore eu feugiat nulla
facilisis at vero eros et accumsan et iusto odio dignissim qui blandit
praesent luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat. Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.%
} % end of abstract

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/about/class> for details.
%% We recommend the 2012 system <http://www.acm.org/about/class/class/2012>
%% For the 2012 system use the ``\CCScatTwelve'' which command takes four arguments.
%% The 1998 system <http://www.acm.org/about/class/class/2012> is still possible
%% For the 1998 system use the ``\CCScat'' which command takes four arguments.
%% In both cases the last two arguments (1998) or last three (2012) can be empty.

\CCScatlist{
  \CCScatTwelve{Human-centered computing}{Visu\-al\-iza\-tion}{Visu\-al\-iza\-tion techniques}{Treemaps};
  \CCScatTwelve{Human-centered computing}{Visu\-al\-iza\-tion}{Visualization design and evaluation methods}{}
}

%\CCScatlist{
  %\CCScat{H.5.2}{User Interfaces}{User Interfaces}{Graphical user interfaces (GUI)}{};
  %\CCScat{H.5.m}{Information Interfaces and Presentation}{Miscellaneous}{}{}
%}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

%% \section{Introduction} %for journal use above \firstsection{..} instead
This template is for papers of VGTC-sponsored conferences which are \emph{\textbf{not}} published in a special issue of TVCG.

\section{Using the Style Template}

\begin{itemize}
\item If you receive compilation errors along the lines of ``\texttt{Package ifpdf Error: Name clash, \textbackslash ifpdf is already defined}'' then please add a new line ``\texttt{\textbackslash let\textbackslash ifpdf\textbackslash relax}'' right after the ``\texttt{\textbackslash documentclass[journal]\{vgtc\}}'' call. Note that your error is due to packages you use that define ``\texttt{\textbackslash ifpdf}'' which is obsolete (the result is that \texttt{\textbackslash ifpdf} is defined twice); these packages should be changed to use ifpdf package instead.
\item The style uses the hyperref package, thus turns references into internal links. We thus recommend to make use of the ``\texttt{\textbackslash autoref\{reference\}}'' call (instead of ``\texttt{Figure\~{}\textbackslash ref\{reference\}}'' or similar) since ``\texttt{\textbackslash autoref\{reference\}}'' turns the entire reference into an internal link, not just the number. Examples: \autoref{fig:sample} and \autoref{tab:vis_papers}.
\item The style automatically looks for image files with the correct extension (eps for regular \LaTeX; pdf, png, and jpg for pdf\LaTeX), in a set of given subfolders (figures/, pictures/, images/). It is thus sufficient to use ``\texttt{\textbackslash includegraphics\{CypressView\}}'' (instead of ``\texttt{\textbackslash includegraphics\{pictures/CypressView.jpg\}}'').
\item For adding hyperlinks and DOIs to the list of references, you can use ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' (instead of ``\texttt{\textbackslash bibliographystyle\{abbrv\}}''). It uses the doi and url fields in a bib\TeX\ entry and turns the entire reference into a link, giving priority to the doi. The doi can be entered with or without the ``\texttt{http://dx.doi.org/}'' url part. See the examples in the bib\TeX\ file and the bibliography at the end of this template.\\[1em]
\textbf{Note 1:} occasionally (for some \LaTeX\ distributions) this hyper-linked bib\TeX\ style may lead to \textbf{compilation errors} (``\texttt{pdfendlink ended up in different nesting level ...}'') if a reference entry is broken across two pages (due to a bug in hyperref). In this case make sure you have the latest version of the hyperref package (i.\,e., update your \LaTeX\ installation/packages) or, alternatively, revert back to ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-narrow\}}'' (at the expense of removing hyperlinks from the bibliography) and try ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' again after some more editing.\\[1em]
\textbf{Note 2:} the ``\texttt{-narrow}'' versions of the bibliography style use the font ``PTSansNarrow-TLF'' for typesetting the DOIs in a compact way. This font needs to be available on your \LaTeX\ system. It is part of the \href{https://www.ctan.org/pkg/paratype}{``paratype'' package}, and many distributions (such as MikTeX) have it automatically installed. If you do not have this package yet and want to use a ``\texttt{-narrow}'' bibliography style then use your \LaTeX\ system's package installer to add it. If this is not possible you can also revert to the respective bibliography styles without the ``\texttt{-narrow}'' in the file name.\\[1em]
DVI-based processes to compile the template apparently cannot handle the different font so, by default, the template file uses the \texttt{abbrv-doi} bibliography style but the compiled PDF shows you the effect of the \texttt{abbrv-doi-hyperref-narrow} style.
\end{itemize}

\section{Bibliography Instructions}

\begin{itemize}
\item Sort all bibliographic entries alphabetically but the last name of the first author. This \LaTeX/bib\TeX\ template takes care of this sorting automatically.
\item Merge multiple references into one; e.\,g., use \cite{Max:1995:OMF,Kitware:2003} (not \cite{Kitware:2003}\cite{Max:1995:OMF}). Within each set of multiple references, the references should be sorted in ascending order. This \LaTeX/bib\TeX\ template takes care of both the merging and the sorting automatically.
\item Verify all data obtained from digital libraries, even ACM's DL and IEEE Xplore  etc.\ are sometimes wrong or incomplete.
\item Do not trust bibliographic data from other services such as Mendeley.com, Google Scholar, or similar; these are even more likely to be incorrect or incomplete.
\item Articles in journal---items to include:
  \begin{itemize}
  \item author names
	\item title
	\item journal name
	\item year
	\item volume
	\item number
	\item month of publication as variable name (i.\,e., \{jan\} for January, etc.; month ranges using \{jan \#\{/\}\# feb\} or \{jan \#\{-{}-\}\# feb\})
  \end{itemize}
\item use journal names in proper style: correct: ``IEEE Transactions on Visualization and Computer Graphics'', incorrect: ``Visualization and Computer Graphics, IEEE Transactions on''
\item Papers in proceedings---items to include:
  \begin{itemize}
  \item author names
	\item title
	\item abbreviated proceedings name: e.\,g., ``Proc.\textbackslash{} CONF\_ACRONYNM'' without the year; example: ``Proc.\textbackslash{} CHI'', ``Proc.\textbackslash{} 3DUI'', ``Proc.\textbackslash{} Eurographics'', ``Proc.\textbackslash{} EuroVis''
	\item year
	\item publisher
	\item town with country of publisher (the town can be abbreviated for well-known towns such as New York or Berlin)
  \end{itemize}
\item article/paper title convention: refrain from using curly brackets, except for acronyms/proper names/words following dashes/question marks etc.; example:
\begin{itemize}
	\item paper ``Marching Cubes: A High Resolution 3D Surface Construction Algorithm''
	\item should be entered as ``\{M\}arching \{C\}ubes: A High Resolution \{3D\} Surface Construction Algorithm'' or  ``\{M\}arching \{C\}ubes: A high resolution \{3D\} surface construction algorithm''
	\item will be typeset as ``Marching Cubes: A high resolution 3D surface construction algorithm''
\end{itemize}
\item for all entries
\begin{itemize}
	\item DOI can be entered in the DOI field as plain DOI number or as DOI url; alternative: a url in the URL field
	\item provide full page ranges AA-{}-BB
\end{itemize}
\item when citing references, do not use the reference as a sentence object; e.\,g., wrong: ``In \cite{Lorensen:1987:MCA} the authors describe \dots'', correct: ``Lorensen and Cline \cite{Lorensen:1987:MCA} describe \dots''
\end{itemize}

\section{Example Section}

Lorem\marginpar{\small You can use the margins for comments while editing the submission, but please remove the marginpar comments for submission.} ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam
voluptua~\cite{Kitware:2003,Max:1995:OMF}. At vero eos et accusam et
justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est.

\section{Exposition}

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat~\cite{Kindlmann:1999:SAG}.

\begin{equation}
\sum_{j=1}^{z} j = \frac{z(z+1)}{2}
\end{equation}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet.

\subsection{Lorem ipsum}

Lorem ipsum dolor sit amet (see \autoref{tab:vis_papers}), consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum.

\begin{table}[tb]
  \caption{VIS/VisWeek accepted/presented papers: 1990--2016.}
  \label{tab:vis_papers}
  \scriptsize%
	\centering%
  \begin{tabu}{%
	r%
	*{7}{c}%
	*{2}{r}%
	}
  \toprule
   year & \rotatebox{90}{Vis/SciVis} &   \rotatebox{90}{SciVis conf} &   \rotatebox{90}{InfoVis} &   \rotatebox{90}{VAST} &   \rotatebox{90}{VAST conf} &   \rotatebox{90}{TVCG @ VIS} &   \rotatebox{90}{CG\&A @ VIS} &   \rotatebox{90}{VIS/VisWeek} \rotatebox{90}{incl. TVCG/CG\&A}   &   \rotatebox{90}{VIS/VisWeek} \rotatebox{90}{w/o TVCG/CG\&A}   \\
  \midrule
	2016 & 30 &   & 37 & 33 & 15 & 23 & 10 & 148 & 115 \\
  2015 & 33 & 9 & 38 & 33 & 14 & 17 & 15 & 159 & 127 \\
  2014 & 34 &   & 45 & 33 & 21 & 20 &   & 153 & 133 \\
  2013 & 31 &   & 38 & 32 &   & 20 &   & 121 & 101 \\
  2012 & 42 &   & 44 & 30 &   & 23 &   & 139 & 116 \\
  2011 & 49 &   & 44 & 26 &   & 20 &   & 139 & 119 \\
  2010 & 48 &   & 35 & 26 &   &   &   & 109 & 109 \\
  2009 & 54 &   & 37 & 26 &   &   &   & 117 & 117 \\
  2008 & 50 &   & 28 & 21 &   &   &   & 99 & 99 \\
  2007 & 56 &   & 27 & 24 &   &   &   & 107 & 107 \\
  2006 & 63 &   & 24 & 26 &   &   &   & 113 & 113 \\
  2005 & 88 &   & 31 &   &   &   &   & 119 & 119 \\
  2004 & 70 &   & 27 &   &   &   &   & 97 & 97 \\
  2003 & 74 &   & 29 &   &   &   &   & 103 & 103 \\
  2002 & 78 &   & 23 &   &   &   &   & 101 & 101 \\
  2001 & 74 &   & 22 &   &   &   &   & 96 & 96 \\
  2000 & 73 &   & 20 &   &   &   &   & 93 & 93 \\
  1999 & 69 &   & 19 &   &   &   &   & 88 & 88 \\
  1998 & 72 &   & 18 &   &   &   &   & 90 & 90 \\
  1997 & 72 &   & 16 &   &   &   &   & 88 & 88 \\
  1996 & 65 &   & 12 &   &   &   &   & 77 & 77 \\
  1995 & 56 &   & 18 &   &   &   &   & 74 & 74 \\
  1994 & 53 &   &   &   &   &   &   & 53 & 53 \\
  1993 & 55 &   &   &   &   &   &   & 55 & 55 \\
  1992 & 53 &   &   &   &   &   &   & 53 & 53 \\
  1991 & 50 &   &   &   &   &   &   & 50 & 50 \\
  1990 & 53 &   &   &   &   &   &   & 53 & 53 \\
  \midrule
  \textbf{sum} & \textbf{1545} & \textbf{9} & \textbf{632} & \textbf{310} & \textbf{50} & \textbf{123} & \textbf{25} & \textbf{2694} & \textbf{2546} \\
  \bottomrule
  \end{tabu}%
\end{table}

\subsection{Mezcal Head}

Lorem ipsum dolor sit amet (see \autoref{fig:sample}), consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet.

\subsubsection{Duis Autem}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est. Lorem
ipsum dolor sit amet.

\begin{figure}[tb]
 \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
 \includegraphics[width=\columnwidth]{paper-count-w-2015-new}
 \caption{A visualization of the 1990--2015 data from \autoref{tab:vis_papers}. The image is from \cite{Isenberg:2017:VMC} and is in the public domain.}
 \label{fig:sample}
\end{figure}

\subsubsection{Ejector Seat Reservation}

Duis autem~\cite{Lorensen:1987:MCA}\footnote{The algorithm behind
Marching Cubes \cite{Lorensen:1987:MCA} had already been
described by Wyvill et al. \cite{Wyvill:1986:DSS} a year
earlier.} vel eum iriure dolor in hendrerit
in vulputate velit esse molestie consequat,\footnote{Footnotes
appear at the bottom of the column.} vel illum dolore eu
feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt
ut laoreet dolore magna aliquam erat volutpat.

\paragraph{Confirmed Ejector Seat Reservation}

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo
consequat~\cite{Nielson:1991:TAD}. Duis autem vel eum iriure dolor in
hendrerit in vulputate velit esse molestie consequat, vel illum dolore
eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi.

\paragraph{Rejected Ejector Seat Reservation}

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie

\section{Conclusion}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum.

%% if specified like this the section will be committed in review mode
\acknowledgments{
The authors wish to thank A, B, and C. This work was supported in part by
a grant from XYZ.}

%\bibliographystyle{abbrv}
\bibliographystyle{abbrv-doi}
%\bibliographystyle{abbrv-doi-narrow}
%\bibliographystyle{abbrv-doi-hyperref}
%\bibliographystyle{abbrv-doi-hyperref-narrow}

\bibliography{paper}
\end{document}

### 2018-09-13 paper.tex
Initial sentences

@@
 
-\title{Global Illumination for Fun and Profit}
+\title{Visual Analysis of Ligand Trajectories in Molecular Dynamics}
 
@@
 %% Abstract section.
-\abstract{Duis autem vel eum iriure dolor in hendrerit in vulputate
-velit esse molestie consequat, vel illum dolore eu feugiat nulla
-facilisis at vero eros et accumsan et iusto odio dignissim qui blandit
-praesent luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat. Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.%
-} % end of abstract
+\abstract{TODO} % end of abstract
 
@@
 %% \section{Introduction} %for journal use above \firstsection{..} instead
-This template is for papers of VGTC-sponsored conferences which are \emph{\textbf{not}} published in a special issue of TVCG.
+TODO
 
-\section{Using the Style Template}
+\section{Related Work}
 
+TODO
+
+\section{Overview}
+
+TODO
 \begin{itemize}
-\item If you receive compilation errors along the lines of ``\texttt{Package ifpdf Error: Name clash, \textbackslash ifpdf is already defined}'' then please add a new line ``\texttt{\textbackslash let\textbackslash ifpdf\textbackslash relax}'' right after the ``\texttt{\textbackslash documentclass[journal]\{vgtc\}}'' call.
-Note that your error is due to packages you use that define ``\texttt{\textbackslash ifpdf}'' which is obsolete (the result is that \texttt{\textbackslash ifpdf} is defined twice); these packages should be changed to use ifpdf package instead.
-\item The style uses the hyperref package, thus turns references into internal links.
-We thus recommend to make use of the ``\texttt{\textbackslash autoref\{reference\}}'' call (instead of ``\texttt{Figure\~{}\textbackslash ref\{reference\}}'' or similar) since ``\texttt{\textbackslash autoref\{reference\}}'' turns the entire reference into an internal link, not just the number.
-Examples: \autoref{fig:sample} and \autoref{tab:vis_papers}.
-\item The style automatically looks for image files with the correct extension (eps for regular \LaTeX; pdf, png, and jpg for pdf\LaTeX), in a set of given subfolders (figures/, pictures/, images/).
-It is thus sufficient to use ``\texttt{\textbackslash includegraphics\{CypressView\}}'' (instead of ``\texttt{\textbackslash includegraphics\{pictures/CypressView.jpg\}}'').
-\item For adding hyperlinks and DOIs to the list of references, you can use ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' (instead of ``\texttt{\textbackslash bibliographystyle\{abbrv\}}'').
-It uses the doi and url fields in a bib\TeX\ entry and turns the entire reference into a link, giving priority to the doi.
-The doi can be entered with or without the ``\texttt{http://dx.doi.org/}'' url part.
-See the examples in the bib\TeX\ file and the bibliography at the end of this template.\\[1em]
-\textbf{Note 1:} occasionally (for some \LaTeX\ distributions) this hyper-linked bib\TeX\ style may lead to \textbf{compilation errors} (``\texttt{pdfendlink ended up in different nesting level ...}'') if a reference entry is broken across two pages (due to a bug in hyperref).
-In this case make sure you have the latest version of the hyperref package (i.\,e., update your \LaTeX\ installation/packages) or, alternatively, revert back to ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-narrow\}}'' (at the expense of removing hyperlinks from the bibliography) and try ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' again after some more editing.\\[1em]
-\textbf{Note 2:} the ``\texttt{-narrow}'' versions of the bibliography style use the font ``PTSansNarrow-TLF'' for typesetting the DOIs in a compact way.
-This font needs to be available on your \LaTeX\ system.
-It is part of the \href{https://www.ctan.org/pkg/paratype}{``paratype'' package}, and many distributions (such as MikTeX) have it automatically installed.
-If you do not have this package yet and want to use a ``\texttt{-narrow}'' bibliography style then use your \LaTeX\ system's package installer to add it.
-If this is not possible you can also revert to the respective bibliography styles without the ``\texttt{-narrow}'' in the file name.\\[1em]
-DVI-based processes to compile the template apparently cannot handle the different font so, by default, the template file uses the \texttt{abbrv-doi} bibliography style but the compiled PDF shows you the effect of the \texttt{abbrv-doi-hyperref-narrow} style.
+  \item We visualize ligand trajectories
+  \item We visualize energetic profiles
 \end{itemize}
 
-\section{Bibliography Instructions}
+From the chemistry point of view, the binding energy of a ligand-substrate complex is an essential measure when evaluating the complex.
+Therefore, we decided to strongly focus on conveying the binding energy when visualizing the trajectories dataset.
+Naturally, this can be achieved by sampling the binding energy along a ligand trajectory and plotting the obtained data to a 2D chart.
+In this manner, we obtain the energetic profile of the trajectory (see Figure ?)
 
+\section{Visual Analysis of Trajectories}
+
+TODO
 \begin{itemize}
-\item Sort all bibliographic entries alphabetically but the last name of the first author. This \LaTeX/bib\TeX\ template takes care of this sorting automatically.
-\item Merge multiple references into one; e.\,g., use \cite{Max:1995:OMF,Kitware:2003} (not \cite{Kitware:2003}\cite{Max:1995:OMF}).
-Within each set of multiple references, the references should be sorted in ascending order.
-This \LaTeX/bib\TeX\ template takes care of both the merging and the sorting automatically.
-\item Verify all data obtained from digital libraries, even ACM's DL and IEEE Xplore  etc.\ are sometimes wrong or incomplete.
-\item Do not trust bibliographic data from other services such as Mendeley.com, Google Scholar, or similar; these are even more likely to be incorrect or incomplete.
-\item Articles in journal---items to include:
-  \begin{itemize}
-  \item author names
-	\item title
-	\item journal name
-	\item year
-	\item volume
-	\item number
-	\item month of publication as variable name (i.\,e., \{jan\} for January, etc.; month ranges using \{jan \#\{/\}\# feb\} or \{jan \#\{-{}-\}\# feb\})
-  \end{itemize}
-\item use journal names in proper style: correct: ``IEEE Transactions on Visualization and Computer Graphics'', incorrect: ``Visualization and Computer Graphics, IEEE Transactions on''
-\item Papers in proceedings---items to include:
-  \begin{itemize}
-  \item author names
-	\item title
-	\item abbreviated proceedings name: e.\,g., ``Proc.\textbackslash{} CONF\_ACRONYNM'' without the year; example: ``Proc.\textbackslash{} CHI'', ``Proc.\textbackslash{} 3DUI'', ``Proc.\textbackslash{} Eurographics'', ``Proc.\textbackslash{} EuroVis''
-	\item year
-	\item publisher
-	\item town with country of publisher (the town can be abbreviated for well-known towns such as New York or Berlin)
-  \end{itemize}
-\item article/paper title convention: refrain from using curly brackets, except for acronyms/proper names/words following dashes/question marks etc.; example:
-\begin{itemize}
-	\item paper ``Marching Cubes: A High Resolution 3D Surface Construction Algorithm''
-	\item should be entered as ``\{M\}arching \{C\}ubes: A High Resolution \{3D\} Surface Construction Algorithm'' or  ``\{M\}arching \{C\}ubes: A high resolution \{3D\} surface construction algorithm''
-	\item will be typeset as ``Marching Cubes: A high resolution 3D surface construction algorithm''
-\end{itemize}
-\item for all entries
-\begin{itemize}
-	\item DOI can be entered in the DOI field as plain DOI number or as DOI url; alternative: a url in the URL field
-	\item provide full page ranges AA-{}-BB
-\end{itemize}
-\item when citing references, do not use the reference as a sentence object; e.\,g., wrong: ``In \cite{Lorensen:1987:MCA} the authors describe \dots'', correct: ``Lorensen and Cline \cite{Lorensen:1987:MCA} describe \dots''
+  \item figures of direct visualization of trajectories, profiles
 \end{itemize}
 
-\section{Example Section}
+\subsection{Small Multiples (Dataset Slicing)}
 
-Lorem\marginpar{\small You can use the margins for comments while editing the submission, but please remove the marginpar comments for submission.} ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam
-voluptua~\cite{Kitware:2003,Max:1995:OMF}. At vero eos et accusam et
-justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est.
+\paragraph{•} Properties -- In order to be able to slice the data, we first define an abstract representation of a ligand trajectory.
+We decided to represent the trajectory using a set of 1D variables that we call trajectory properties.
+Each trajectory property is defined such that it describes a specific characterstic of the trajectory.
+We define the following properties:
+\begin{itemize}
+  \item \textbf{Minimum Distance to Active Site} --
+  \item \textbf{Area Under Curve} --
+  \item \textbf{?Energy Variance} --
+  \item \textbf{Structure RMSD} --
+  \item \textbf{Ligand Conformation Variance} --
+\end{itemize}
 
-\section{Exposition}
+\paragraph{•} Trends
 
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat~\cite{Kindlmann:1999:SAG}.
+\paragraph{•} Zooming
 
-\begin{equation}
-\sum_{j=1}^{z} j = \frac{z(z+1)}{2}
-\end{equation}
+\subsection{3D Visualization}
 
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet.
+\paragraph{•} Spatial Aggregation
 
-\subsection{Lorem ipsum}
+\section{Evaluation}
 
-Lorem ipsum dolor sit amet (see \autoref{tab:vis_papers}), consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum.
+TODO
+\begin{itemize}
+  \item LinBwt vs. LinB32 (W177 closes main tunnel)
+\end{itemize}
 
-\begin{table}[tb]
-  \caption{VIS/VisWeek accepted/presented papers: 1990--2016.}
-  \label{tab:vis_papers}
-  \scriptsize%
-	\centering%
-  \begin{tabu}{%
-	r%
-	*{7}{c}%
-	*{2}{r}%
-	}
-  \toprule
-   year & \rotatebox{90}{Vis/SciVis} &   \rotatebox{90}{SciVis conf} &   \rotatebox{90}{InfoVis} &   \rotatebox{90}{VAST} &   \rotatebox{90}{VAST conf} &   \rotatebox{90}{TVCG @ VIS} &   \rotatebox{90}{CG\&A @ VIS} &   \rotatebox{90}{VIS/VisWeek} \rotatebox{90}{incl.
-TVCG/CG\&A}   &   \rotatebox{90}{VIS/VisWeek} \rotatebox{90}{w/o TVCG/CG\&A}   \\
-  \midrule
-	2016 & 30 &   & 37 & 33 & 15 & 23 & 10 & 148 & 115 \\
-  2015 & 33 & 9 & 38 & 33 & 14 & 17 & 15 & 159 & 127 \\
-  2014 & 34 &   & 45 & 33 & 21 & 20 &   & 153 & 133 \\
-  2013 & 31 &   & 38 & 32 &   & 20 &   & 121 & 101 \\
-  2012 & 42 &   & 44 & 30 &   & 23 &   & 139 & 116 \\
-  2011 & 49 &   & 44 & 26 &   & 20 &   & 139 & 119 \\
-  2010 & 48 &   & 35 & 26 &   &   &   & 109 & 109 \\
-  2009 & 54 &   & 37 & 26 &   &   &   & 117 & 117 \\
-  2008 & 50 &   & 28 & 21 &   &   &   & 99 & 99 \\
-  2007 & 56 &   & 27 & 24 &   &   &   & 107 & 107 \\
-  2006 & 63 &   & 24 & 26 &   &   &   & 113 & 113 \\
-  2005 & 88 &   & 31 &   &   &   &   & 119 & 119 \\
-  2004 & 70 &   & 27 &   &   &   &   & 97 & 97 \\
-  2003 & 74 &   & 29 &   &   &   &   & 103 & 103 \\
-  2002 & 78 &   & 23 &   &   &   &   & 101 & 101 \\
-  2001 & 74 &   & 22 &   &   &   &   & 96 & 96 \\
-  2000 & 73 &   & 20 &   &   &   &   & 93 & 93 \\
-  1999 & 69 &   & 19 &   &   &   &   & 88 & 88 \\
-  1998 & 72 &   & 18 &   &   &   &   & 90 & 90 \\
-  1997 & 72 &   & 16 &   &   &   &   & 88 & 88 \\
-  1996 & 65 &   & 12 &   &   &   &   & 77 & 77 \\
-  1995 & 56 &   & 18 &   &   &   &   & 74 & 74 \\
-  1994 & 53 &   &   &   &   &   &   & 53 & 53 \\
-  1993 & 55 &   &   &   &   &   &   & 55 & 55 \\
-  1992 & 53 &   &   &   &   &   &   & 53 & 53 \\
-  1991 & 50 &   &   &   &   &   &   & 50 & 50 \\
-  1990 & 53 &   &   &   &   &   &   & 53 & 53 \\
-  \midrule
-  \textbf{sum} & \textbf{1545} & \textbf{9} & \textbf{632} & \textbf{310} & \textbf{50} & \textbf{123} & \textbf{25} & \textbf{2694} & \textbf{2546} \\
-  \bottomrule
-  \end{tabu}%
-\end{table}
-
-\subsection{Mezcal Head}
-
-Lorem ipsum dolor sit amet (see \autoref{fig:sample}), consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet.
-
-\subsubsection{Duis Autem}
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est. Lorem
-ipsum dolor sit amet.
-
-\begin{figure}[tb]
- \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
- \includegraphics[width=\columnwidth]{paper-count-w-2015-new}
- \caption{A visualization of the 1990--2015 data from \autoref{tab:vis_papers}. The image is from \cite{Isenberg:2017:VMC} and is in the public domain.}
- \label{fig:sample}
-\end{figure}
-
-\subsubsection{Ejector Seat Reservation}
-
-Duis autem~\cite{Lorensen:1987:MCA}\footnote{The algorithm behind
-Marching Cubes \cite{Lorensen:1987:MCA} had already been
-described by Wyvill et al. \cite{Wyvill:1986:DSS} a year
-earlier.} vel eum iriure dolor in hendrerit
-in vulputate velit esse molestie consequat,\footnote{Footnotes
-appear at the bottom of the column.} vel illum dolore eu
-feugiat nulla facilisis at vero eros et accumsan et iusto odio
-dignissim qui blandit praesent luptatum zzril delenit augue duis
-dolore te feugait nulla facilisi. Lorem ipsum dolor sit amet,
-consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt
-ut laoreet dolore magna aliquam erat volutpat.
-
-\paragraph{Confirmed Ejector Seat Reservation}
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo
-consequat~\cite{Nielson:1991:TAD}. Duis autem vel eum iriure dolor in
-hendrerit in vulputate velit esse molestie consequat, vel illum dolore
-eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
-dignissim qui blandit praesent luptatum zzril delenit augue duis
-dolore te feugait nulla facilisi.
-
-\paragraph{Rejected Ejector Seat Reservation}
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
+\section{Future Work}
 
@@
 
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum.
-
 %% if specified like this the section will be committed in review mode
-\acknowledgments{
-The authors wish to thank A, B, and C. This work was supported in part by
-a grant from XYZ.}
+\acknowledgments{TODO This work was supported in part by a grant from XYZ.}
 

### 2018-09-14 paper.tex
prototype figures

@@
 
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\columnwidth]{trajectories}
+  \caption{TODO replace, caption}
+  \label{fig:trajectories}
+\end{figure}
+
 From the chemistry point of view, the binding energy of a ligand-substrate complex is an essential measure when evaluating the complex.
@@
 Naturally, this can be achieved by sampling the binding energy along a ligand trajectory and plotting the obtained data to a 2D chart.
-In this manner, we obtain the energetic profile of the trajectory (see Figure ?)
+In this manner, we obtain an abstraction of the trajectory in terms of the ligand binding energy profile (see Figure~\ref{fig:trajectory}).
+Further, the energy profiles of the trajectories can be visualized within one 2D plot using superimposition.
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.5\columnwidth]{trajectory}\\
+  \vspace{1em}
+  \includegraphics[width=0.8\columnwidth]{trajectory-energy}
+  \caption{TODO figure quality, caption}
+  \label{fig:trajectory}
+\end{figure}
+
+However, the superimposed visualization of energetic profiles does not scale well and becomes uninformative when visualizing hundreds (and more) trajectories (see Figure ?).
+Particularly, there arise two issues for the users of such visualization. First, the main energetic trends become overshadowed, and second it is not possible to navigate to individual trajectories.
+In order to overcome these issues, we propose a visualization system which employs well established visualization concepts in an interactive manner such that a ligand trajectories dataset can be explored based on the user needs.
+Namely, we reduce the amount of energetic profiles visualized in a chart by binning the data by neighborhood and presenting it using small multiples approach.
+Regarding the binning, we aim for as general as possible usage and therefore we let the user to control the data neighborhood measure.
+We describe the approach in detail in Section 4.?.
 
@@
 
-\section{Future Work}
+\section{Discussion (+ Future Work)}
+
+TODO
+\begin{itemize}
+  \item can be used also for CaverDock
+\end{itemize}
 

### 2018-09-14 paper.tex
Started overview

@@
 %% \section{Introduction} %for journal use above \firstsection{..} instead
-TODO
+\begin{itemize}
+  \item Motivation
+  \item Ligand binding (Method + Trajectory data)
+  \item Contribution
+\end{itemize}
 
@@
 
+\begin{itemize}
+  \item Trajectories visualization
+  \item Differences -- ligand MD trajectory, waters
+  \item Missing?
+\end{itemize}
+
+\section{Overview}
+
 TODO
 
-\section{Overview}
+\subsection{Ligand Binding Trajectories}
+
+TODO Jak moc se budeme rozepisovat? ENDTODO
+When designing the proposed visualization system, we used ligand trajectories data that was computed using our own approach to the ligand binding simulation.
+Our simulation method is inspired by robotics, and it exploits a motion planning algorithm~\cite{lavalle2006planning} in order to navigate the ligand through the void space of a molecule.
+Specifically, we implemented a modification of the Rapidly-exploring random tree (RRT)~\cite{lavalle1998rapidly} algorithm.
+We modified the original algorithm such that it does not sample the void space uniformly.
+Instead, we first compute the Voronoi diagram which represents the void space whithin a molecule.
+In order to do this, we represent the molecule using van der Waals spheres of its atoms and we compute the aditively weighted Voronoi diagram of atom spheres.
+Then, we employ the Voronoi diagram during the sample generation phase of the RRT algorithm such that we bias the sample position towards the Voronoi vertices [VCBM2018? or else?].
+Our modification is based on the assumption that it is more probable to find a ligand transport path near the largest voids within a molecule.
+This assumption is also utilized by the state-of-the-art molecular tunnel detection algorithms~\cite{yaffe2008molaxis, chovancova2012caver, sehnal2013mole}.
+Although we implemented our own ligand binding simulation algorithm, the proposed visualization system is not restricted to it.
+On the contrary, we designed the visualization with other ligand binding simulation tools in mind (see Section~\ref{sec:discussion}).
+
+TODO Autodock Vina?
+
+TODO describe visualization input data
+TODO describe direct 3D visualization?
 
@@
 \section{Discussion (+ Future Work)}
+\label{sec:discussion}
 

### 2018-09-14 paper.tex
Overview mini changes

@@
 In order to overcome these issues, we propose a visualization system which employs well established visualization concepts in an interactive manner such that a ligand trajectories dataset can be explored based on the user needs.
-Namely, we reduce the amount of energetic profiles visualized in a chart by binning the data by neighborhood and presenting it using small multiples approach.
-Regarding the binning, we aim for as general as possible usage and therefore we let the user to control the data neighborhood measure.
-We describe the approach in detail in Section 4.?.
+Namely, we reduce the amount of energetic profiles visualized in a chart by binning the data and presenting it using small multiples~\cite{tufte1990envisioning} approach.
+We bin the trajectories according to their location within a paremetric space which we define based on 1D variables that describe the trajectories (see Section~\ref{sec:multiples}).
+We aim for as general as possible usage and therefore we let the user to control the variables that define the parametric space.
 
@@
 \subsection{Small Multiples (Dataset Slicing)}
+\label{sec:multiples}
 
@@
 \begin{itemize}
-  \item can be used also for CaverDock
+  \item can be used also for CaverDock, MoMA-LigPath (Vina run needed)
 \end{itemize}

### 2018-09-15 paper.tex
Overview continued

@@
 
-TODO
+We propose an interactive visualization system which enables chemists to explore datasets that emerge from ligand-receptor binding simulations.
+Such a dataset can be described as a set of trajectories where each trajectory describes a ligand path to the interior of a molecule.
+Mostly, the trajectories are discrete to some extent, e.g., in terms of movement.
+The discrete nature follows directly from the current simulation methods, e.g., the Molecular dynamics (MD) simulation.
+Therefore, it is sufficiently general to describe a ligand binding trajectory as a set of samples where each sample represents a ligand-receptor complex at a given time.
+
+TODO describe general trajectory
+
+TODO describe our specific trajectory
 
@@
 TODO Jak moc se budeme rozepisovat? ENDTODO
+
 When designing the proposed visualization system, we used ligand trajectories data that was computed using our own approach to the ligand binding simulation.
@@
 This assumption is also utilized by the state-of-the-art molecular tunnel detection algorithms~\cite{yaffe2008molaxis, chovancova2012caver, sehnal2013mole}.
-Although we implemented our own ligand binding simulation algorithm, the proposed visualization system is not restricted to it.
+In this manner, we compute a non-continuous trajectory for a ligand to the interior of a molecule through its void space.
+Finally, we evaluate the trajectories in terms of the ligand binding energy using the AutoDock Vina~\cite{trott2010autodock} molecular docking tool.
+For this, we treat each sample of a trajectory as a ligand-receptor complex for which we compute its binding energy, i.e., AutoDock Vina's binding affinity.
+
+Although we implemented our own ligand binding simulation algorithm, the proposed visualization system is not limited to it.
 On the contrary, we designed the visualization with other ligand binding simulation tools in mind (see Section~\ref{sec:discussion}).
-
-TODO Autodock Vina?
 
@@
 \end{figure}
+
+\subsection{Binding Energy Profile}
 

### 2018-09-24 paper.tex
Case study #1 draft

@@
 %% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
-\graphicspath{{figures/}{pictures/}{images/}{./}} % where to search for the images
+\graphicspath{{figures/}{pictures/}{images/}{figures/study/}{./}} % where to search for the images
 
@@
 
+Ve Fig 3 je videt, jak v LinBwt ligand prochazi skoro vzdycky.
+Naopak ve Fig 4 je videt, ze v LinB32 ligand projde malokdy.
+To odpovida laboratornim datum o snizene aktivite po mutaci (W177).
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\columnwidth]{cs01-linbwt-all}
+  \caption{TODO}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\columnwidth]{cs01-linb32-all}
+  \caption{TODO}
+\end{figure}
+
+Po prozkoumani horni skupiny u Fig 4 je videt, ze se ligandy zastavily pri pruchodu okolo mutovane W177.
+Narust energie je videt na Fig 6.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\columnwidth]{cs01-linb32-sel-failed-3d}
+  \caption{TODO}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\columnwidth]{cs01-linb32-sel-failed}
+  \caption{TODO}
+\end{figure}
+
+Kdyz se podivame na trajektorie, ktere byly uspesne na LinBwt a LinB32, tak to vypada takto. Fig 7 vs. Fig 8.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\columnwidth]{cs01-linb32-sel-success}
+  \caption{TODO}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\columnwidth]{cs01-linbwt-sel-02}
+  \caption{TODO}
+\end{figure}
+
+Pri pohledu ve 3D to na LinB32 vypada takto (Fig 9).
+Mutace snizila pruchodnost tunelu p1.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\columnwidth]{cs01-linb32-sel-success-3d}
+  \caption{TODO}
+\end{figure}
+
+Dalsi use case by mel ukazat, ze nekdy, kdyz ligand projde na LinB32, konformace molekuly se zmenila.
+
 \section{Discussion (+ Future Work)}

### 2018-09-26 paper.tex
Few comments on motivation

@@
 %% \section{Introduction} %for journal use above \firstsection{..} instead
+
+Motivace:
 \begin{itemize}
-  \item Motivation
+  \item Proteiny jsou essential to all living...
+  \item Specificky druh jsou enzymy... metabolismus?
+  \item Dnes chemici bezne zkoumaji funkci enzymu pomoci modelovani, predevsim pomoci molekularni simulace.
+  \item Vysledky in-sillico experimentu jsou objemne a je bezne je analyzovat predem navrzenou metodou s zadnou ci malou znalosti zkoumane simulace [https://loschmidt.chemi.muni.cz/peg/publications/conformational-changes-allow-processing-of-bulky-substrates-by-a-haloalkane-dehalogenase-with-a-small-and-buried-active-site/].
+  \item Cilem naseho vizualizacniho nastroje je umoznit interaktivne "zjistit, co se v datech deje", s malou ci zadnou predstavou o metode jejich zpracovani.
+  Myslim to tak, ze metodu zpracovani si chemik bude moci navrhnout/zpresnit zaroven se samotnou analyzou dat.
+\end{itemize}
+
+Dalsi:
+\begin{itemize}
   \item Ligand binding (Method + Trajectory data)

### 2018-09-27 paper.tex

--- 2018-09-24 paper.tex
@@
 \usepackage{booktabs}                  % only used for the table example
+\usepackage{color}
 %% We encourage the use of mathptmx for consistent usage of times font
@@
 %% \section{Introduction} %for journal use above \firstsection{..} instead
+
+Proteins are essential parts of living organisms. 
+Understanding the function of the organisms and their malfunction therefore starts with deep understanding of protein structure and function.
+The reactivity of proteins with other molecules is the key property which enables to create new molecules as products of chemical reactions. 
+These molecules can serve for different purposes, e.g., can serve as the basis of new drugs. 
+In protein engineering, the researchers are focusing on the properties of proteins themselves and the chemical reactions are designed in order to modify these properties. 
+An example can be changing the stability of proteins in normal conditions, outside the lab environment~\cite{koudelakova}, or increasing the protein activity towards other molecules~\cite{pavlova}.
+
+Enzymes are forming a specific group of proteins, which accelerate chemical reactions. 
+Their have an irreplaceable function in organisms as the enzyme catalysis plays a role in almost all metabolic processes in cells.
+The important property of enzymes is that their reaction rate can be increased by lowering their activation energy.
+Therefore, studying the energetic profile of the reaction of a small molecule, called ligand, with the enzyme gives the biochemists the information about the feasibility of this reaction to happen.
+
+In other words, when the biochemists are simulating (i.e., modeling) the reaction between the enzyme and ligand, they are using in-silico methods for generating  possible trajectories of the ligand to the enzyme deeply buried site, called the active site.
+In this specific site the chemical reaction between the enzyme and ligand can undergo. 
+The existing computational methods are generating large ensembles of such trajectories which have to be further explored by the biochemists who are assessing their relevance.
+To do this, they have to focus in different properties of the trajectories, including its energetic profile. 
+Currently the exploration of possible trajectories is performed by a specific method designed by the biochemists especially for the particular case~\cite{pmid29858243}.
+However, a general approach is still missing.
+The straightforward solution is designing visualization and visual analysis methods for exploration of trajectories and their properties, enabling the user to interactively filter them according to interesting geometric and physico-chemical aspects.
+In this paper, we are presenting a novel tool for visual analysis of large ensembles of possible ligand trajectories, whose design was driven by the actual needs of the protein engineers.
+They were also tightly collaborating on the design of the tool, in order to address the needs as much as possible. 
+Our tool consists of several linked views enabling the user to explore crucial properties of large ensembles of ligand trajectories, containing thousands of them. 
+These can be generated by different simulation algorithms and tools, such as \textcolor{red}{TODO cite caverdock and Vojta's method}.
+%~\cite{caverdock,vojta}.
+
+The ultimate goal of our tool is to enable the biochemists to get insight into the relevance of trajectories even when they do not have any apriori knowledge about the trajectories and how to design a method for their processing in a traditionally used manner.
+
+\textcolor{red}{TODO finish that}
+
+%Motivace:
+%\begin{itemize}
+%  \item Proteiny jsou essential to all living...
+%  \item Specificky druh jsou enzymy... metabolismus?
+%  \item Dnes chemici bezne zkoumaji funkci enzymu pomoci modelovani, predevsim pomoci molekularni simulace.
+%  \item Vysledky in-sillico experimentu jsou objemne a je bezne je analyzovat predem navrzenou metodou s zadnou ci malou znalosti zkoumane simulace [https://loschmidt.chemi.muni.cz/peg/publications/conformational-changes-allow-processing-of-bulky-substrates-by-a-haloalkane-dehalogenase-with-a-small-and-buried-active-site/].
+%  \item Cilem naseho vizualizacniho nastroje je umoznit interaktivne "zjistit, co se v datech deje", s malou ci zadnou predstavou o metode jejich zpracovani.
+%  Myslim to tak, ze metodu zpracovani si chemik bude moci navrhnout/zpresnit zaroven se samotnou analyzou dat.
+%\end{itemize}
+
+Dalsi:
 \begin{itemize}
-  \item Motivation
   \item Ligand binding (Method + Trajectory data)

### 2018-09-27 paper.tex
intro finished and abstract

@@
 %% Abstract section.
-\abstract{TODO} % end of abstract
+\abstract{
+Exploration of protein reactivity with other molecules plays a crucial role in many biochemical research disciplines, including protein engineering and drug design.
+In many cases, a small molecule (ligand) reacts with the protein in its deeply buried active site. 
+As the ligand has to be transported to this site from the protein outer environment, it is crucial to find a feasible trajectory, which the ligand can follow.
+There are several in-silico methods, enabling to calculate potential trajectories. 
+They usually produce large ensembles of trajectories which have to be further explored by the biochemists who are trying to assess their relevance. 
+This decision is based on many parameters of the trajectories which have to be taken into account.
+This makes the exploration very demanding as the biochemists have to analyze the multidimensional data about trajectories.
+Without proper visual support, this process is very complicated and time demanding.
+In this paper, we introduce our tool for visual analysis of ligand trajectories and their various properties.
+The tool has been designed in tight collaboration with the domain experts in order to address their needs as much as possible.
+It is composed of several linked views, enabling interactive exploration and filtering of trajectories.
+The tool consists of a set of 2D views and the 3D view.
+It utilizes the concept of a scatterplot matrix for communicating different properties of the trajectories.
+\textcolor{red}{TODO describe in one sentence the other views as well}
+The usefulness of the tool is demonstrated on two case studies, conducted by the collaborating protein engineers.
+} % end of abstract
 
@@
 
-\textcolor{red}{TODO finish that}
+In the following, we are first discussing techniques and methods related to our approach and then we describe our tool in detail. 
+This is followed by the demonstration of the usefulness of our tool on two case studies, performed on real scenarios and data from the latest research of our collaborating protein engineers.
 

### 2018-09-27 paper.tex
updates abstract

@@
 The tool consists of a set of 2D views and the 3D view.
-It utilizes the concept of a scatterplot matrix for communicating different properties of the trajectories.
-\textcolor{red}{TODO describe in one sentence the other views as well}
+It utilizes the concept of a small multiples for communicating different properties of the trajectories.
+Through variety of interaction techniques the system enables users to create and store selections that can be compared and evaluated later.
+In both 2D and 3D views, users can either observer the individual trajectories or choose to aggregate the information into a functional boxplot or a density visualization.
+%It utilizes the concept of a scatterplot matrix for communicating different properties of the trajectories.
 The usefulness of the tool is demonstrated on two case studies, conducted by the collaborating protein engineers.

### 2018-09-27 paper.tex
correction
Revert "correction"
This reverts commit 4119588cf6b5c7eb4c8de44464d2675d5b4d8f52.

@@
 In many cases, a small molecule (ligand) reacts with the protein in its deeply buried active site. 
-As the ligand has to be transported to this site from the protein outer environment, it is crucial to find a feasible trajectory, which the ligand can follow.
-There are several in-silico methods, enabling to calculate potential trajectories. 
-They usually produce large ensembles of trajectories which have to be further explored by the biochemists who are trying to assess their relevance. 
-This decision is based on many parameters of the trajectories which have to be taken into account.
+As the ligand has to be transported to this site from the protein outer environment, it is crucial to study trajectories, which the ligand can follow.
+There are several in-silico methods, enabling to calculate potential trajectories in a protein model.
+When these methods are applied in connection with molecular dynamics simulation, usually large ensembles of trajectories are produced.
+These trajectories have to be further explored by the biochemists who are trying to assess their relevance and gain insight into the protein function.
+During this process many parameters of the trajectories have to be taken into account.
 This makes the exploration very demanding as the biochemists have to analyze the multidimensional data about trajectories.
 Without proper visual support, this process is very complicated and time demanding.
-In this paper, we introduce our tool for visual analysis of ligand trajectories and their various properties.
+In this paper, we introduce a novel tool for visual analysis of ligand trajectories and their various properties.
 The tool has been designed in tight collaboration with the domain experts in order to address their needs as much as possible.
-It is composed of several linked views, enabling interactive exploration and filtering of trajectories.
-The tool consists of a set of 2D views and the 3D view.
-It utilizes the concept of a small multiples for communicating different properties of the trajectories.
+It is composed of several 2D and 3D linked views, enabling interactive exploration and filtering of trajectories.
+%The tool consists of a set of 2D views and the 3D view.
+It utilizes the concept of a small multiples for communicating the properties of the trajectories.
 Through variety of interaction techniques the system enables users to create and store selections that can be compared and evaluated later.

### 2018-09-27 paper.tex
Updates the abstract based on the suggestions given by Grammarly.

@@
 These trajectories have to be further explored by the biochemists who are trying to assess their relevance and gain insight into the protein function.
-During this process many parameters of the trajectories have to be taken into account.
+During this process, many parameters of the trajectories have to be taken into account.
 This makes the exploration very demanding as the biochemists have to analyze the multidimensional data about trajectories.
@@
 %The tool consists of a set of 2D views and the 3D view.
-It utilizes the concept of a small multiples for communicating the properties of the trajectories.
-Through variety of interaction techniques the system enables users to create and store selections that can be compared and evaluated later.
+It utilizes the concept of small multiples for communicating the properties of the trajectories.
+Through a variety of interaction techniques, the system enables users to create and store selections that can be compared and evaluated later.
 In both 2D and 3D views, users can either observer the individual trajectories or choose to aggregate the information into a functional boxplot or a density visualization.
 %It utilizes the concept of a scatterplot matrix for communicating different properties of the trajectories.
-The usefulness of the tool is demonstrated on two case studies, conducted by the collaborating protein engineers.
+The usefulness of the tool is demonstrated in two case studies, conducted by the collaborating protein engineers.
 } % end of abstract

### 2018-09-28 paper.tex
Added authors

@@
 %% Author and Affiliation (multiple authors with multiple affiliations)
-\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ %
-        \scriptsize Starbucks Research %
-\and Ed Grimley\thanks{e-mail: ed.grimley@aol.com}\\ %
-     \scriptsize Grimley Widgets, Inc. %
-\and Martha Stewart\thanks{e-mail: martha.stewart@marthastewart.com}\\ %
-     \parbox{1.4in}{\scriptsize \centering Martha Stewart Enterprises \\ Microsoft Research}}
+\author{Adam Jur\v{c}\'{i}k\thanks{e-mail: xjurc@fi.muni.cz}\\ %
+        \scriptsize Masaryk University %
+\and Katar\'{i}na Furmanov\'{a}\\ %
+     \scriptsize Masaryk University %
+\and Jan By\v{s}ka\\ %
+     \parbox{1.4in}{\scriptsize \centering University of Bergen \\ Masaryk University} %
+\and Vojt\v{e}ch Von\'{a}sek\\ %
+     \scriptsize Czech Technical University %
+\vspace{\baselineskip} %
+\and Ond\v{r}ej V\'{a}vra\\ %
+     \scriptsize Masaryk University %
+\and Pavol Ulbrich\\ %
+     \scriptsize Masaryk University
+\and Helwig Hauser\\ %
+     \scriptsize University of Bergen %
+\and Barbora Kozl\'{i}kov\'{a}\\ %
+     \scriptsize Masaryk University}
 

### 2018-09-28 paper.tex
Updates abstract base ond the low level comments from Helwig (TODO: implement the high level comments as well).

--- 2018-09-27 paper.tex
@@
 \abstract{
-Exploration of protein reactivity with other molecules plays a crucial role in many biochemical research disciplines, including protein engineering and drug design.
-In many cases, a small molecule (ligand) reacts with the protein in its deeply buried active site. 
-As the ligand has to be transported to this site from the protein outer environment, it is crucial to study trajectories, which the ligand can follow.
-There are several in-silico methods, enabling to calculate potential trajectories in a protein model.
+Exploration of protein reactions with other molecules plays a crucial role in biochemical research, including protein engineering and drug design.
+In many cases, a small molecule (ligand) reacts with the protein at a deeply buried active site. 
+As the ligand has to get to this site from the outer environment, it is crucial to study trajectories of ligand motion.
+There are several in-silico methods to calculate potential trajectories.
 When these methods are applied in connection with molecular dynamics simulation, usually large ensembles of trajectories are produced.
-These trajectories have to be further explored by the biochemists who are trying to assess their relevance and gain insight into the protein function.
+These trajectories have to be explored by biochemists who are trying to assess their relevance and gain insight into the protein function.
 During this process, many parameters of the trajectories have to be taken into account.
@@
 Without proper visual support, this process is very complicated and time demanding.
-In this paper, we introduce a novel tool for visual analysis of ligand trajectories and their various properties.
-The tool has been designed in tight collaboration with the domain experts in order to address their needs as much as possible.
-It is composed of several 2D and 3D linked views, enabling interactive exploration and filtering of trajectories.
+In this paper, we introduce a novel tool for the interactive visual exploration and analysis of ligand trajectories and their various properties.
+The tool has been designed in tight collaboration with domain experts in order to address their needs as much as possible.
+It is composed of several linked 2D and 3D views, enabling the interactive exploration and filtering of trajectories.
 %The tool consists of a set of 2D views and the 3D view.
-It utilizes the concept of small multiples for communicating the properties of the trajectories.
+Small multiples are used for communicating the properties of the trajectories.
 Through a variety of interaction techniques, the system enables users to create and store selections that can be compared and evaluated later.

### 2018-09-28 paper.tex
new abstract we created based on Helwigs comments

@@
 \abstract{
-Exploration of protein reactions with other molecules plays a crucial role in biochemical research, including protein engineering and drug design.
-In many cases, a small molecule (ligand) reacts with the protein at a deeply buried active site. 
-As the ligand has to get to this site from the outer environment, it is crucial to study trajectories of ligand motion.
-There are several in-silico methods to calculate potential trajectories.
-When these methods are applied in connection with molecular dynamics simulation, usually large ensembles of trajectories are produced.
-These trajectories have to be explored by biochemists who are trying to assess their relevance and gain insight into the protein function.
-During this process, many parameters of the trajectories have to be taken into account.
-This makes the exploration very demanding as the biochemists have to analyze the multidimensional data about trajectories.
-Without proper visual support, this process is very complicated and time demanding.
-In this paper, we introduce a novel tool for the interactive visual exploration and analysis of ligand trajectories and their various properties.
-The tool has been designed in tight collaboration with domain experts in order to address their needs as much as possible.
-It is composed of several linked 2D and 3D views, enabling the interactive exploration and filtering of trajectories.
-%The tool consists of a set of 2D views and the 3D view.
-Small multiples are used for communicating the properties of the trajectories.
-Through a variety of interaction techniques, the system enables users to create and store selections that can be compared and evaluated later.
-In both 2D and 3D views, users can either observer the individual trajectories or choose to aggregate the information into a functional boxplot or a density visualization.
-%It utilizes the concept of a scatterplot matrix for communicating different properties of the trajectories.
+In many cases, protein reactions with other small molecules (ligands) occur at a deeply buried active site. 
+When studying these types of reactions it is crucial for biochemists to examine trajectories of ligand motion.
+These trajectories are currently predicted with in-silico methods that usually produce large ensembles of possible trajectories. 
+In this paper, we propose a novel approach to visual exploration and analysis of these large datasets allowing domain experts to assess the relevance of individual trajectories based on their properties.
+The proposed solution is composed of multiple linked 2D and 3D views, enabling the interactive exploration and filtering of trajectories in an informed way.
+In the workflow we describe we focus on practical aspects of the interactive visual analysis specific to the ligand trajectories data.
+We adopted the well-proven approach of small multiples which allows us to break down the large dataset into smaller chunks that are easier to operate.
+Here we describe how the drill down technique can be used to create and store selections of trajectories with desired properties which allows us to compare multiple datasets.
+In both 2D and 3D views, biochemists can either observer the individual trajectories or choose to aggregate the information into a functional boxplot or a density visualization.
+The tool was developed in tight collaboration with domain experts in order to address their needs as much as possible.
 The usefulness of the tool is demonstrated in two case studies, conducted by the collaborating protein engineers.
+%Exploration of protein reactions with other molecules plays a crucial role in biochemical research, including protein engineering and drug design.
+%In many cases, a small molecule (ligand) reacts with the protein at a deeply buried active site. 
+%As the ligand has to get to this site from the outer environment, it is crucial to study trajectories of ligand motion.
+%There are several in-silico methods to calculate potential trajectories.
+%When these methods are applied in connection with molecular dynamics simulation, usually large ensembles of trajectories are produced.
+%These trajectories have to be explored by biochemists who are trying to assess their relevance and gain insight into the protein function.
+%During this process, many parameters of the trajectories have to be taken into account.
+%This makes the exploration very demanding as the biochemists have to analyze the multidimensional data about trajectories.
+%Without proper visual support, this process is very complicated and time demanding.
+%In this paper, we introduce a novel tool for the interactive visual exploration and analysis of ligand trajectories and their various properties.
+%The tool has been designed in tight collaboration with domain experts in order to address their needs as much as possible.
+%It is composed of several linked 2D and 3D views, enabling the interactive exploration and filtering of trajectories.
+%Small multiples are used for communicating the properties of the trajectories.
+%Through a variety of interaction techniques, the system enables users to create and store selections that can be compared and evaluated later.
+%In both 2D and 3D views, users can either observer the individual trajectories or choose to aggregate the information into a functional boxplot or a density visualization.
+%The usefulness of the tool is demonstrated in two case studies, conducted by the collaborating protein engineers.
 } % end of abstract

### 2018-09-28 paper.tex
adds a list of possible related work papers

@@
 \begin{itemize}
+	\item General molecular vis:
+		\begin{itemize}
+			\item Visual Analysis of Biomolecular Cavities: State of the Art
+			\item Visualization of biomolecular structures: State of the art revisited.
+		\end{itemize}
   	\item Trajectories visualization
-  \item Differences -- ligand MD trajectory, waters
+  		\begin{itemize}
+  			\item Interactive Visual Analysis of Families of Function Graphs
+  			\item Curve Boxplot: Generalization of Boxplot for Ensembles of Curves
+  			\item Streamline Variability Plots for Characterizing the Uncertainty in Vector Field Ensembles
+  			\item {\color{red} simplifikace ve 3D, maybe not related} SimpliFly: A Methodology for Simplification and Thematic Enhancement of Trajectories
+  			\item Generalized box-plot for root growth ensembles
+  		\end{itemize}
+  	\item relevant for the modulate 3D visualization
+  		\begin{itemize}
+  			\item Game On, Science - How Video Game Technology May Help Biologists Tackle Visualization Challenges
+  		\end{itemize}
+  	\item relevant for density visualization
+  		\begin{itemize}
+  			\item Spatial Generalization and Aggregation of Massive Movement Data
+  			\item Space–time density of trajectories: exploring spatio-temporal patterns in movement data
+  		\end{itemize}
+  	\item relevant for selection by density
+  		\begin{itemize}
+  			\item Visualizing Dynamical Systems near Critical Points
+  			\item Strategies for Interactive Exploration of 3D Flow Using Evenly-Spaced Illuminated Streamlines
+  		\end{itemize}
+  	\item Trajectories in molecular context
+  		\begin{itemize}
+  			\item Watergate: Visual Exploration of Water Trajectories in Protein Dynamics
+			\item Interactive Exploration of Ligand Transportation through Protein Tunnels
+  			\item MolPathFinder: Interactive Multi-Dimensional Path Filtering of Molecular Dynamics Simulation Data
+  			\item {\color{red}not completely relevant} Methodologies for the Analysis of Instantaneous Lipid Diffusion in MD Simulations of Large Membrane Systems
+  			\item {\color{red}maybe not relevant} Heterogeneity-based Guidance for Exploring Multiscale Data in Systems Biology
+			\item {\color{red}water trajectories - 3D} Visual Abstractions of Solvent Pathlines near Protein Cavities
+			\item {\color{red}water trajectories - 3D - domain paper} Tracking the Flow of Water through Photosystem II Using Molecular
+			\item {\color{red}water trajectories - 3D - domain paper - visualization by pymol} AQUA-DUCT: a ligands tracking tool Dynamics and Streamline Tracing
+  		\end{itemize}
   \item Missing?

### 2018-09-28 paper.tex
helwig's comments to the v2 of the abstract

@@
 When studying these types of reactions it is crucial for biochemists to examine trajectories of ligand motion.
-These trajectories are currently predicted with in-silico methods that usually produce large ensembles of possible trajectories. 
-In this paper, we propose a novel approach to visual exploration and analysis of these large datasets allowing domain experts to assess the relevance of individual trajectories based on their properties.
+These trajectories are predicted with in-silico methods that produce large ensembles of possible trajectories.
+In this paper, we propose a novel approach to the interactive visual exploration and analysis of large sets of trajectories enabling domain experts to assess the relevance of individual trajectories based on their properties.
 The proposed solution is composed of multiple linked 2D and 3D views, enabling the interactive exploration and filtering of trajectories in an informed way.
-In the workflow we describe we focus on practical aspects of the interactive visual analysis specific to the ligand trajectories data.
-We adopted the well-proven approach of small multiples which allows us to break down the large dataset into smaller chunks that are easier to operate.
-Here we describe how the drill down technique can be used to create and store selections of trajectories with desired properties which allows us to compare multiple datasets.
-In both 2D and 3D views, biochemists can either observer the individual trajectories or choose to aggregate the information into a functional boxplot or a density visualization.
-The tool was developed in tight collaboration with domain experts in order to address their needs as much as possible.
-The usefulness of the tool is demonstrated in two case studies, conducted by the collaborating protein engineers.
+In the workflow, we focus on the practical aspects of the interactive visual analysis specific to ligand trajectories.
+We adopt and adapt small multiples to resolve an overly large number of trajectories into smaller chunks that are easier to analyze.
+We describe how drill-down techniques can be used to create and store selections of trajectories with desired properties, enabling the comparison of multiple datasets.
+In appropriately crafted 2D and 3D views, biochemists can either observe individual trajectories or choose to aggregate the information into a functional boxplot or density visualization.
+Our solution is based on a tight collaboration with domain experts in order to address their needs as much as possible.
+The usefulness of this novel approach is demonstrated by two case studies, conducted by the collaborating protein engineers.
+
 %Exploration of protein reactions with other molecules plays a crucial role in biochemical research, including protein engineering and drug design.

### 2018-09-28 paper.tex
Adams feedback on abstract

@@
 These trajectories are predicted with in-silico methods that produce large ensembles of possible trajectories.
-In this paper, we propose a novel approach to the interactive visual exploration and analysis of large sets of trajectories enabling domain experts to assess the relevance of individual trajectories based on their properties.
+In this paper, we propose a novel approach to the interactive visual exploration and analysis of large sets of ligand trajectories enabling domain experts to understand protein function based on the trajectory properties.
 The proposed solution is composed of multiple linked 2D and 3D views, enabling the interactive exploration and filtering of trajectories in an informed way.
